---
title: "Exam#2"
author: "Thakur Prasad Ghimire"
date: "11/18/2021"
output: word_document
---
 









 1.  Hypothesis test 
Null hypothesis: people with various educational qualifications who are full or parttime have different fraction offered health insurance.
2. (20 points) Using a subset of the data (so you can do this question without R), the following coefficients are estimated in a logit model where employer offering health insurance depends on $Age$, $Age^2$, $Female$, and interaction of $Female$ with the age terms.

Variable       | Coeff    | Std Error
-------------- | -------  | ----------
$Age$          | -0.019   | 0.0018
$Age^2$        | 0.00002  | 0.000002
$Female$       | -0.470   | 0.1206
$Age*Female$   | 0.0082   | 0.0026
$Age^2*Female$ | -0.00001 | 0.000003
$Constant$     | 1.84     | 0.09

  a. What is the predicted probability that a 30-year-old person (not a female) will have a job with benefits?
  
  
Log(p/((1-p) ))=-0.019*30+0.00002*(30)^2-0.470*0+0.0082*30*1-0.00001*(30)^2*1+1.84=-1.993
p=0.0233



  b. What is the predicted probability that a 30-year-old female person will have a job with benefits?
  
  
  	Log(p/((1-p) ))=-0.019*30+0.00002*(30)^2-0.470*1+0.082*1-0.00001*1+1.84=2.002
	p=0.9772
  c. Describe and explain the impact of the squared age terms in the model, for male and female.


Explore likelihood of people working at a job that offers health insurance.

#3
```{r}
library(tidyverse)
library(ggplot2)
library(stats)
library(dplyr)
library(randomForest)
library(corrplot)
load("C:/Homework EcoB2000/Exam#2/NHIS2020_data (2).RData")

```

Choose a subgroup of the sample to consider and provide summary statistics of that subgroup.
#Choose the black race sub group

```{r}
data_NHIS_subgrp =  data_NHIS[data_NHIS$RACEA=="Black",]
summary(data_NHIS_subgrp)
```

Explain why this subgroup is interesting: This subgroup is mostly located in the south region, and most of the black people have high school diploma as their highest education level, and most of them have health insurance.

Form a hypothesis test about an interesting variable, explore whether your chosen subgroup differs from the rest of sample. Please provide both a p-value for the hypothesis test and a confidence interval. Write a short paragraph explaining the test (carefully noting what is the null hypothesis) and explaining the results of that test.
The interesting variable is HOURSWRK. Null hypothesis: Hours of work is not correlated with someones age.

Now we will test this hypothesis on both the sample and the rest of the data to see if it holds.

With a linear probability model (OLS), can you find relevant information to predict an interesting outcome? How good is the model? Discuss. Include hypothesis tests about individual or joint coefficient values. Include some predicted values.
Lets predict the hours worked by a black person in USA using age, sex, and citizenship. The null hypothesis is that these predictors have little or no effect on the number of hours worked by a black person in the US.

``{r}
```{r}
logit_model = glm(HOURSWRK ~ AGE + SEX  + CITIZEN,data = data_NHIS_subgrp)
summary(logit_model)
```
The null hypothesis as mentioned earlier that Hours of work is not affected by these variables (age, sex, citizen).
The logit model performs the same way as the OLS model because the same variables like female sex, US citizen are the ones that are significant here too.


#Correlation test on the sub group
```{r}
cor.test(data_NHIS_subgrp$HOURSWRK,data_NHIS_subgrp$AGE)
```
The test shows that there was no correlation between age and hours of work for black people.

#Rest of the data
```{r}
rest_df = data_NHIS[data_NHIS$RACEA!="Black",]
```

#Correlation test on the rest of the data
```{r}
cor.test(rest_df$HOURSWRK,rest_df$HOURSWRK)
```
The test on the rest of the data showed that there was correlation between hours of work and age. Therefore the sub group differed with the rest of the data here.

- With a linear probability model (OLS), can you find relevant information to predict an interesting outcome? How good is the model? Discuss. Include hypothesis tests about individual or joint coefficient values. Include some predicted values.

Lets predict the hours worked by a black person in USA using age, sex, and citizenship. The null hypothesis is that these predictors have little or no effect on the number of hours worked by a black person in the US.  

```{r}
LPM = lm(HOURSWRK ~ AGE + SEX  + CITIZEN,data = data_NHIS_subgrp)
summary(LPM)
```


```{r}

logit_model = glm(HOURSWRK ~ AGE + SEX  + CITIZEN,data = data_NHIS_subgrp)
summary(logit_model)
```
The null hypothesis as mentioned earlier that Hours of work is not affected by these variables (age, sex, citizen).
The logit model performs the same way as the OLS model because the same variables like female sex, US citizen are the ones that are significant here too.


Thakur Prasad ghimire 
City College of New york

